Democracy, Public Support, and Measurement Uncertainty

### 2021-01-01 paper/dcpo_demsupport.Rmd
replace "Here" with "In this letter" to stress to reviewers that this is a short article with 4k wordcount limit
revise introduction further

---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: svm-latex-ms2.tex
title: "Democracy, Public Support, and Measurement Uncertainty"
thanks: "The authors contributed equally to this work.  Corresponding author: [yuehu@mail.tsinghua.edu.cn](mailto:yuehu@mail.tsinghua.edu.cn).  Current version: `r format(Sys.time(), '%B %d, %Y')`."
author:
- name: Yuehong 'Cassandra' Tai
  affiliation: University of Iowa
- name: Yue Hu
  affiliation: Tsinghua University
- name: Frederick Solt
  affiliation: University of Iowa
abstract: "Do democratic regimes depend on public support to avoid backsliding? Does public support, in turn, respond thermostatically to changes in democracy? Two prominent recent studies (Claassen 2020a, Claassen 2020b) have reinvigorated the classic hypothesis on the positive relationship between public support for democracy and regime survival--and challenged its reciprocal counterpart--by using a Bayesian latent variable approach to measure mass democratic support from cross-national survey data. But such approaches come with concomitant measurement uncertainty, and neither study incorporated this uncertainty into its analyses. Here we correctly take this uncertainty in account and show that there is no support for the conclusion of either study. We then work to minimize the measurement uncertainty in public support by bringing additional survey data and a superior model of public opinion. Even with these improvements, however, our analyses fail to yield evidence in support of either hypothesis. [142/150 words]"
keywords: ""
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
spacing: double
bibliography: \dummy{`r file.path(getwd(), list.files(getwd(), ".bib$", recursive = TRUE))`}
# csl: https://raw.githubusercontent.com/citation-style-language/styles/master/american-political-science-association.csl
biblio-style: apsr
citecolor: black
linkcolor: black
endnote: no
header-includes:
      - \usepackage{array}
      - \usepackage{caption}
      - \usepackage{graphicx}
      - \usepackage{siunitx}
      - \usepackage{colortbl}
      - \usepackage{multirow}
      - \usepackage{hhline}
      - \usepackage{calc}
      - \usepackage{tabularx}
      - \usepackage{threeparttable}
      - \usepackage{wrapfig}
---
# Introduction

@@

One important reason for these mixed results is the difficulty in measuring democratic support over time and across many countries.
Public support for democracy cannot be directly observed, and its incorrect measurement will limit inferences about the relationships between public opinion and institutional development.
Further, the survey data available across countries and over time on support for democracy--or indeed most topics in public opinion---are sparse and incomparable, greatly hindering the broadly comparative research needed to reach sound conclusions.
Recently, a few pioneering studies have sought to overcome the hurdle of sparse and incomparable data and advance broadly comparative research by developing latent variable measurement models of public opinion [see @Caughey2019; @Claassen2019; @Solt2020a].
A pair of prominent recent works took advantage of this latent variable approach to measure dynamic public democratic support for 135 countries for up to 29 years and to then assess its consequences for democratic change.
It concluded that mass support had a positive impact on democratic change, especially the endurance of democracy.

The models employed in this study's analyses, though, do not account for measurement uncertainty.
Because they are unobserved, latent variables are inherently accompanied by a quantity of measurement uncertainty.
To leave this uncertainty unacknowledged is to make the implausible assumption that the latent variables are measured perfectly, an assumption which distorts both statistical and substantive inference [@Crabtree2015; @juhl_measurement_2019].

This paper reexamines the classic claim about public democratic attitude and democratic changes thoroughly through both sophisticated methods and bigger data.
In addition to incorporating uncertainty, we employed a more sophisticated measurement, the Dynamic Comparative Public Opinion (DCPO) model developed by @Solt2020a, to estimate democratic support for a larger dataset with 144 countries for up to 33 years between 1988 and 2020.
Surprisingly, our study revealed that the significant relationship between public support and democratic change disappeared once uncertainty was included in the replication analysis of @Claassen2020a model with his original data and the DCPO analysis with more data.
To be cautious, we then also reexamined @Claassen2020b analyses by following the same operations. 
Again, when uncertainty was incorporated, the significant relationship between democratic changes and public democratic attitude disappeared.

The disappearing relationship might be caused by theoretical, methodological reasons, or both.
Methodologically, the existing survey questions might not be able to capture mass democratic attitude due to a shortage of indicators [Solt, 2021??] and the ineffectiveness of indicators in measuring the multi-faceted democratic support [@wuttke2020have] and behaviorally democratic commitments [@graham_democracy_2020; @Carey2020; @McCoy2020].
Substantive explanations might also account for the insignificant relationship: the elite-driven democratic backsliding [@weingast_political_1997; @levitsky_how_2018], the lack of a consensus on democratic erosion between the public and experts [@carey_searching_2019], and the difference between the public support in abstract and public commitments in context [@bartels_ethnic_2020; @graham_democracy_2020; @McCoy2020; @prothro_fundamental_1960; @van2017myth
].

@@

# Model Specification 
To examine the influence of empirical uncertainty, we conducted a three-step test. 
The first step we made is to reproduce @Claassen2020a.
Then applying the method he described in the article, we replicated his IRT-based measurement of democratic mood and maintained the entire distribution of the predicted outcome instead of aggregating it to a single point estimate.
Utilizing the estimated distribution, we involved uncertainty in the analysis of the influence of democratic mood on institutional democracy cross-nationally.
In the third step, we applied the DCPO measurement and added one-third more data to the analysis to reduce potential effects of measurement biases and data spareness.
We compared the outcomes of the three steps and showed as below that uncertainty is a critical factor in the studies of mass democratic attitude, which may alter the analytic conclusions. 
We further showed that accounting for uncertainty is equivalently important when the latent-variable estimates play as a dependent variable in a study by replicating @Claassen2020b.

>>>>

---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: svm-latex-ms2.tex
title: "Democracy, Public Support, and Measurement Uncertainty"
thanks: "The authors contributed equally to this work.  Corresponding author: [yuehu@mail.tsinghua.edu.cn](mailto:yuehu@mail.tsinghua.edu.cn).  Current version: `r format(Sys.time(), '%B %d, %Y')`."
author:
- name: Yuehong 'Cassandra' Tai
  affiliation: University of Iowa
- name: Yue Hu
  affiliation: Tsinghua University
- name: Frederick Solt
  affiliation: University of Iowa
abstract: "Do democratic regimes depend on public support to avoid backsliding? Does public support, in turn, respond thermostatically to changes in democracy? Two prominent recent studies (Claassen 2020a,b) have reinvigorated the classic hypothesis on the positive relationship between public support for democracy and regime survival--and challenged its reciprocal counterpart--by using a Bayesian latent variable approach to measure mass democratic support from cross-national survey data. But such approaches come with concomitant measurement uncertainty, and neither study incorporated this uncertainty into its analyses. In this letter, we correctly take this uncertainty in account and show that there is no support for the conclusion of either study. We then work to minimize the measurement uncertainty in public support by bringing additional survey data and a superior model of public opinion. Even with these improvements, however, our analyses fail to yield evidence in support of either hypothesis. [141/150 words]"
keywords: ""
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
spacing: double
bibliography: \dummy{`r file.path(getwd(), list.files(getwd(), ".bib$", recursive = TRUE))`}
# csl: https://raw.githubusercontent.com/citation-style-language/styles/master/american-political-science-association.csl
biblio-style: apsr
citecolor: black
linkcolor: black
endnote: no
header-includes:
      - \usepackage{array}
      - \usepackage{caption}
      - \usepackage{graphicx}
      - \usepackage{siunitx}
      - \usepackage{colortbl}
      - \usepackage{multirow}
      - \usepackage{hhline}
      - \usepackage{calc}
      - \usepackage{tabularx}
      - \usepackage{threeparttable}
      - \usepackage{wrapfig}
---
# Introduction

@@

One important reason for these mixed results is the difficulty in measuring democratic support over time and across many countries.
Public support for democracy cannot be directly observed, and its incorrect measurement will limit inferences about the relationships between public opinion and institutional development.
Further, the survey data available across countries and over time on support for democracy--or indeed most topics in public opinion---are sparse and incomparable, greatly hindering broadly comparative research.
Recently, a few pioneering studies have sought to overcome the hurdle of sparse and incomparable data by developing latent variable measurement models of public opinion [see @Caughey2019; @Claassen2019; @Solt2020a].
A pair of prominent recent works took advantage of this latent variable approach to measure democratic support for over one hundred countries for up to nearly three decades and to then assess, respectively, its consequences for and roots in democratic change [@Claassen2020a; @Claassen2020b].
These works concluded, first and supporting the classic argument, that mass support had a positive impact on democratic change, especially the endurance of democracy [@Claassen2020a, 127-130], and, second and directly contrary to it, democratic change has a thermostatic effect on public support, that is, that rather than generating their own support, deepening democracies provoke a backlash [@Claassen2020b, 46-50].

The models employed in these studies' analyses, though, do not account for uncertainty in their measurement of democratic support.
Because they are unobserved, latent variables are inherently accompanied by a quantity of measurement uncertainty.
To leave this uncertainty unacknowledged is to make the implausible assumption that the latent variables are measured perfectly, an assumption which distorts both statistical and substantive inference [@Crabtree2015; @juhl_measurement_2019].

In this letter, we reexamine the classic arguments about support for democracy and democratic change tested in these two pieces while correcting this oversight.
In addition to incorporating the measurement uncertainty, we also sought to reduce it by employing a superior model of public opinion, the Dynamic Comparative Public Opinion (DCPO) model developed by @Solt2020a, and more data to estimate democratic support for 144 countries for up to 33 years between 1988 and 2020.
Our analyses reveal that the significant relationships between public support and democratic change disappear once measurement uncertainty is taken into account, both in replications of @Claassen2020a and @Claassen2020b with the studies' original data and in our extension DCPO analyses that incorporate additional data.
That is, once measurement uncertainty is accounted for, there is no empirical support for either claim: declining democratic support does not appear important to subsequent democratic backsliding, and changes in democracy do not spur a thermostatic response in democratic support.

These null results might be caused by theoretical, methodological reasons, or both.
Methodologically, the existing survey questions might not be able to capture mass democratic attitude due to a shortage of indicators [Solt, 2021??] and the ineffectiveness of indicators in measuring the multi-faceted democratic support [@wuttke2020have] and behaviorally democratic commitments [@graham_democracy_2020; @Carey2020; @McCoy2020].
Substantive explanations might also account for the insignificant relationship: the elite-driven democratic backsliding [@weingast_political_1997; @levitsky_how_2018], the lack of a consensus on democratic erosion between the public and experts [@carey_searching_2019], and the difference between the public support in abstract and public commitments in context [@bartels_ethnic_2020; @graham_democracy_2020; @McCoy2020; @prothro_fundamental_1960; @van2017myth
].

@@

# Model Specification 
To examine the influence of empirical uncertainty, we conducted a three-step test. 
The first step we made is to reproduce @Claassen2020a.
Then, applying the method he described in the article, we replicated his IRT-based measurement of democratic mood and maintained the entire distribution of the predicted outcome instead of aggregating it to a single point estimate.
Utilizing the estimated distribution, we involved uncertainty in the analysis of the influence of democratic mood on institutional democracy cross-nationally.
In the third step, we applied the DCPO measurement and added one-third more data to the analysis to reduce potential effects of measurement biases and data spareness.
We compared the outcomes of the three steps and showed as below that uncertainty is a critical factor in the studies of mass democratic attitude, which may alter the analytic conclusions. 
We further showed that accounting for uncertainty is equivalently important when the latent-variable estimates play as a dependent variable in a study by replicating @Claassen2020b.
